Gammapy: A Python package for gamma-ray astronomy

### 2023-03-21 src/ms-review.tex
add magic ex
Signed-off-by: Atreyee Sinha <asinha@ucm.es>

%% Context: development of open software
In the past decade observing the \gammaray sky has transitioned from a niche in the field of
particle physics to an established branch of astronomy, completing the view of the
sky in high energies. At the same time \textit{Python}
has become extremely popular as a scientific programming language,
in particular in the field of data sciences. This success is
mostly attributed to the simple and easy to learn syntax, the ability to act as
a \enquote{glue} language between different programming languages and last but not least
the rich ecosystem of packages and its open and supportive community \citep{Momcheva2015}.

@@

%% \gammapy: concept and goals
%%% Question: starting date? Is is relevant, what is the real start?
The \gammapy project was started following the model of \astropy, with the objective of building a common
software library for very high-energy \gammaray data analysis \citep{gammapy_2015}. 
The core of the idea is illustrated in Figure~\ref{fig:big_picture}. The various \gammaray instruments
can export their data to a common data format (\gadf) and then these data can be combined and
analysed using a common software library.
The \gammapy package is an independent community-developed software project, it has been selected to be the core library for the Science Analysis tools of CTAO
but also involves contributors associated to other instruments.
The \gammapy package is built on the scientific Python ecosystem: it uses \numpy~\citep{numpy} for n-dimensional data
structures, \scipy~\citep{2020SciPy-NMeth} for numerical algorithms, \astropy~\citep{astropy} for
astronomy-specific functionality, \iminuit~\citep{iminuit} for numerical minimisation
and \matplotlib~\citep{matplotlib} for visualization.

%% CN: what about mentioning the joint Crab? or is it mentioned later?
With the public availability of the \gadf format specifications and the
\gammapy package, some experiments started to make limited subsets of
their \gammaray data publicly available for testing and validating
\gammapy. For example, the \hess collaboration released a limited test
dataset (about 50 hours of observations taken between 2004 and 2008)
based on the \gadf DL3 format \citep{HESS_DR1}. This data release served
as a basis for validation of open analysis tools, including \gammapy 
\cite[see e.g.][]{Mohrmann2019}. The \hawc collaboration also released
a limited test dataset of the Crab Nebula, which was used to validate
the \gammapy package in \cite{Olivera2022}.
The increased availability of public data that followed the definition of
a common data format, and the development of \gammapy as a community-driven open software,
led the way toward a more open science in the very-high-energy \gammaray Astronomy domain.
In future CTAO will be an open observatory committed to follow the FAIR
(Findable, Accessible, Interoperable and Reusable) principles \citep{FAIR16, FAIR22} 
that define the key requirements for open science.

>>>>

%% Context: development of open software
In the past decade, observing the \gammaray sky has transitioned from a niche in the field of
particle physics to an established branch of astronomy, completing the view of the
sky in high energies. At the same time \textit{Python}
has become extremely popular as a scientific programming language,
in particular in the field of data sciences. This success is
mostly attributed to the simple and easy to learn syntax, the ability to act as
a \enquote{glue} language between different programming languages and last but not least
the rich ecosystem of packages and its open and supportive community \citep{Momcheva2015}.

@@

%% \gammapy: concept and goals
%%% Question: starting date? Is is relevant, what is the real start?
The \gammapy project was started following the model of \astropy, with the objective of building a common
software library for very high-energy \gammaray data analysis \citep{gammapy_2015}. 
The core of the idea is illustrated in Figure~\ref{fig:big_picture}. The various \gammaray instruments
can export their data to a common data format (\gadf) and then these data can be combined and
analysed using a common software library.
The \gammapy package is an independent community-developed software project. It has been selected to be the core library for the Science Analysis tools of CTAO
but also involves contributors associated to other instruments.
The \gammapy package is built on the scientific Python ecosystem: it uses \numpy~\citep{numpy} for n-dimensional data
structures, \scipy~\citep{2020SciPy-NMeth} for numerical algorithms, \astropy~\citep{astropy} for
astronomy-specific functionality, \iminuit~\citep{iminuit} for numerical minimisation
and \matplotlib~\citep{matplotlib} for visualization.

They consist of two
observations of $20\,{\rm min}$ each, chosen from the dataset used to estimate
the performance of the upgraded stereo system~\citep{magic_performance} and
already included in~\cite{joint_crab}.

With the public availability of the \gadf format specifications and the
\gammapy package, some experiments started to make limited subsets of
their \gammaray data publicly available for testing and validating
\gammapy. For example, the \hess collaboration released a limited test
dataset (about 50 hours of observations taken between 2004 and 2008)
based on the \gadf DL3 format \citep{HESS_DR1}. This data release served
as a basis for validation of open analysis tools, including \gammapy 
\cite[see e.g.][]{Mohrmann2019}. Two observations of the Crab nebula have been
 released by the \magic collaboration~\citep{magic_performance}, using which a
joint analysis across different gamma-ray instruments was first demonstrated in ~\cite{joint_crab}.
The \hawc collaboration also released
a limited test dataset of the Crab Nebula, which was used to validate
the \gammapy package in \cite{Olivera2022}.
The increased availability of public data that followed the definition of
a common data format, and the development of \gammapy as a community-driven open software,
led the way toward a more open science in the very-high-energy \gammaray Astronomy domain.
In future CTAO will be an open observatory committed to follow the FAIR
(Findable, Accessible, Interoperable and Reusable) principles \citep{FAIR16, FAIR22} 
that define the key requirements for open science.
